---
title: "Seurat - Combining Several 10X Runs, Osteosarcoma "
author: "Sulev Koks"
date: "2024-04-14"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this project, we will combine six 10X Osteosarcoma datasets from the GEO accession GSE162454.

This work is based on:

Hao et al. Dictionary learning for integrative, multimodal and scalable single-cell analysis.
  Nature Biotechnology (2023) [Seurat V5]

To start, we read in the data and create two `Seurat` objects.

```{r}
# install.packages("devtools")
devtools::install_github("immunogenomics/presto")
```

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(presto)
```
```{r load data}
load("OS_sce1_6_filesUNFILTERED.RData")
load("OS1_6merged.RData")
```

```{r}
sce1f<- CreateSeuratObject(counts = assay(sce1, "counts"), project = "OSscRNA1")
sce1f
sce2f<- CreateSeuratObject(counts = assay(sce2, "counts"), project = "OSscRNA2")
sce2f
sce3f<- CreateSeuratObject(counts = assay(sce3, "counts"), project = "OSscRNA3")
sce3f
sce4f<- CreateSeuratObject(counts = assay(sce4, "counts"), project = "OSscRNA4")
sce4f
sce5f<- CreateSeuratObject(counts = assay(sce5, "counts"), project = "OSscRNA5")
sce5f
sce6f<- CreateSeuratObject(counts = assay(sce6, "counts"), project = "OSscRNA6")
sce6f
```

## Filter out mitochondria and outliers

Keep MT % below five and features between 300 and 4500.

```{r}
sce1f[["percent.mt"]] <- PercentageFeatureSet(sce1f, pattern = "^MT-")
sce2f[["percent.mt"]] <- PercentageFeatureSet(sce2f, pattern = "^MT-")
sce3f[["percent.mt"]] <- PercentageFeatureSet(sce3f, pattern = "^MT-")
sce4f[["percent.mt"]] <- PercentageFeatureSet(sce4f, pattern = "^MT-")
sce5f[["percent.mt"]] <- PercentageFeatureSet(sce5f, pattern = "^MT-")
sce6f[["percent.mt"]] <- PercentageFeatureSet(sce6f, pattern = "^MT-")

```
```{r}
sce1f <- subset(sce1f, subset = nFeature_RNA > 300 & nFeature_RNA < 4500 & percent.mt < 5)
sce2f <- subset(sce2f, subset = nFeature_RNA > 300 & nFeature_RNA < 4500 & percent.mt < 5)
sce3f <- subset(sce3f, subset = nFeature_RNA > 300 & nFeature_RNA < 4500 & percent.mt < 5)
sce4f <- subset(sce4f, subset = nFeature_RNA > 300 & nFeature_RNA < 4500 & percent.mt < 5)
sce5f <- subset(sce5f, subset = nFeature_RNA > 300 & nFeature_RNA < 4500 & percent.mt < 5)
sce6f <- subset(sce6f, subset = nFeature_RNA > 300 & nFeature_RNA < 4500 & percent.mt < 5)
```

Normalise if you like

```{r}
sce1fn <- NormalizeData(sce1f)
sce2fn <- NormalizeData(sce2f)
sce3fn <- NormalizeData(sce3f)
sce4fn <- NormalizeData(sce4f)
sce5fn <- NormalizeData(sce5f)
sce6fn <- NormalizeData(sce6f)
```

To merge more than two `Seurat` objects, simply pass a vector of multiple `Seurat` objects to the `y` parameter for `merge`.


```{r}
OS1_6merged <- merge(sce1f, y = c(sce2f, sce3f, sce4f, sce5f, sce6f), add.cell.ids = c('OS1', 'OS2', 'OS3', 'OS4', 'OS5', 'OS6'), project = 'OSscRNAmerged')
OS1_6merged
```
```{r}
head(colnames(OS1_6merged))
tail(colnames(OS1_6merged))
unique(sapply(X = strsplit(colnames(OS1_6merged), split = '_'), FUN = '[', 1))
table(OS1_6merged$orig.ident)
```

```{r}
Layers(OS1_6merged)
```

```{r}
LayerData(OS1_6merged)[1:10, 1:15]
```

```{r save.times, include = TRUE}
write.csv(x = t(as.data.frame(all_times)), file = "merge_vignette_times.csv")
```

```{r}
sessionInfo()
```
